Proportional multi-state multiple-cohort life table model

### 2018-05-01 MSLT/MSLTdoc.Rmd
Tidy up code

#Global options
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = 'latex')
knitr::opts_chunk$set(fig.cap = TRUE, fig.align = "center", fig.pos = "H", fig.scap = TRUE)

#The work in this document requires the following packages (code, rmarkdown and citing packages):
require(dplyr)
require(tidyverse)
require(knitr)
require(kableExtra)
require(citr)
require(gridExtra)
require(cowplot)
require(ggpubr)
```

@@

```{r, echo=FALSE}
pmslt_inputs <- tribble(
~Input, ~Source, ~Comments, 
"Life table", "Synthetic population per sex and age group", "Age grouping in life table to match synthetic population", 
"Life table", "Synthetic population per sex and one-year age group", "If one year age group is not avabilable it can be derive using interpolation from age groups data", 
"Life table", "Global Burden of Disease (GBD) study per one-year age group and sex", "GBD data is in five-year age groups, interpolation to derive one-year age groups", 
"Disease life table", "GBD data for prevalence, incidence and mortality and DISMOD II", "Two step process. First obtain disease and population data from GBD. Second, use Dismod II to derive internally consistent estimates for incidence and case fatality (PMSLT disease life table iputs)", 
"Disease life table", "Derive from disease prevalence and years lived with disability from GBD", "Adjustments for comorbidities in later years of life to be applied"
)

@@

```{r, echo=FALSE}
pmslt_inputs <- tribble(
~Tab, ~Selection,
"Base", "Single", "Location", "Case study city", "Year", "Latest available", "Context", "Cause", "Age", "Under 5, 5 to 9, 10 to 14, 15 to 19, 20 to 24, 25 to 29, 30 to 34, 35 to 39, 40 to 49, 50 to 54, 55 to 59, 60 to 64, 65 to 69, 70 to 74, 75 to 79, 80 to 84, 89 to 89, 90 to 95, 95 plus", "Metric", "Number, Rate", "Measure", "Deaths, YLDs, Prevalence, Incidence", "Sex", "Male, Female", "Cause", "Total All causes, ischemic heart disease, etc")

kable(pmslt_inputs,"latex", booktabs = T, caption = "Global burden of disease data") %>%
 kable_styling(latex_options = c("striped", "hold position"),
               full_width = F)%>%
  column_spec(2, width = "30em") %>%
  column_spec(3, width = "30em") %>%
  row_spec(0, bold = T)
```

@@

4) Read GBD data
```{r, echo=TRUE}
GBDdata <- read.csv("data/UK/gbd2016.csv", stringsAsFactors = F)
```

@@

5) Change all upper cases to lower case and delete () from variables. 
```{r, echo=TRUE}
GBDdata <- mutate_all(GBDdata, funs(tolower))
GBDdata$measure[GBDdata$measure=="ylds (years lived with disability)"] <- "ylds"
```

6) Create age categories index in GBDdata (mid age, to match cohort running), total of 20 age groups. These are the age cohorts to simulate. 
```{r, echo=TRUE}
# 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97
GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 7
GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 12
GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 17
GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 22
GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 27
GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 32
GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 37
GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 42
GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 47
GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 52
GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 57
GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 62
GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 67
GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 72
GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 77
GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 82
GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 87
GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 92
GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97
```

7) Create age and sex categories to obtain population numbers. Population numbers from GBD are used in Dismod II. For the Life table (Figure 3), the numbers may be from the synthetic population. For now, the Life table is set up with population numbers derived from the GBD data. 
```{r, echo=TRUE}
GBDdata$sex_age_cat <- paste(GBDdata$sex,GBDdata$age_cat, sep = "_"  )
```
8) Convert string variables to numeric to do calculations.
```{r, echo=TRUE}
GBDdata$val <- as.numeric(as.character(GBDdata$val))
```
9) Generate population numbers for Greater London in a new data frame ("GBD_population"). Note that there is data for England as well, which is used in a separate rmarkdown document (GBDCompare). Population numbers are derived from rates per 100,000 and total numbers of cases. 
```{r, echo=TRUE}
GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", 
metric == "rate" | metric == "number" ) %>% select(metric, age_cat, sex, val, 
sex_age_cat, location)
```
10) Generate population numbers from given number of cases and rates per 100,000 people. 
```{r, echo=TRUE}
 for (i in 1:nrow(GBD_population)) {
    if (GBD_population$metric[i] == "number") {
      GBD_population$five_year_population[i] <- GBD_population$val[i] * 100000/
      GBD_population$val[i + 2]}     
    else {GBD_population$five_year_population[i] <- NA}
 }
```
11) Remove rows with zero
```{r, echo=TRUE}
GBD_population <- GBD_population[!is.na(GBD_population$five_year_population),]
```
12) Keep relevant variables
```{r, echo=TRUE}
GBD_population <- filter(GBD_population) %>% 
select(sex_age_cat, sex, age_cat, five_year_population, location)
```
13) Create data frames for Greater London to be later used for: a) interpolation of rates and b) PMLT cohorts. 
```{r, echo=TRUE}
GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% 
select(sex_age_cat, age_cat, sex, five_year_population, location, sex_age_cat)
```
12) Check population total numbers
```{r, echo=TRUE}
GreaterLondon <- sum(GBD_population_GL$five_year_population)
```

13) Generate data frames for Greater London with per person rates (per 100,000 in original data).  
```{r, echo=TRUE}
GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% 
select(measure, location, sex, age, metric, cause, val, age_cat, sex_age_cat)
GBDGL$one_rate <- GBDGL$val/100000
```

@@

# Name data frame
assign(paste(sex_index, measure_index, "interpolated", sep = "_"), interpolated)
# plot(interpolated$age,interpolated$rate)

##Do graph with another layer for original rates in age groups for comparison purposes.
p <- ggplot(data = interpolated,mapping = aes(age, interpolated[,1])) +
              geom_line(aes(color = "Interpolated")) +
              geom_point(
              data = data, 
              mapping = aes(age_cat, one_rate, color = "Original")) +
              labs(colour="",x="Age",y="Rates", sep = " ") +
              labs (title = paste("Rates", sex_index, "all cause", 
              measure_index, sep = " "), size=14) +
              theme_classic() +
              theme (plot.title = element_text(hjust = 0.5))
              print(p)
  }
}

@@

```{r, echo=TRUE}
## Create empty data frame
input_life_table <- as.data.frame(matrix(0, ncol=2, nrow = 202))
## Give variables names (mx=total mortality rate, 
## yldsx = total years lived with disability rates )
names(input_life_table) <- c("age", "sex")
## Populate life table: Age, 0 to 100
input_life_table[1:101, 1] <- c(0:100)
input_life_table[102:202, 1] <- c(0:100)
## Populate life table: sex, male, female
input_life_table[1:101, 2] <- "male"
input_life_table[102:202, 2] <- "female"
## Create variable age_sex to match with population data. 
input_life_table$sex_age_cat <- paste(input_life_table$sex,input_life_table$age,
                                sep = "_"  )
##Populate life table: five_year_pop.

input_life_table <- merge(input_life_table, select(GBD_population_GL, c(sex_age_cat, five_year_population)), by = 'sex_age_cat', all = TRUE) 
##Populate life table: mortality rates (from interpolated rates)
## Males mortality rates. NEED MATCHING NAMES
input_life_table <- merge(input_life_table, select(male_deaths_interpolated , c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
input_life_table <- merge(input_life_table, select(female_deaths_interpolated ,c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
input_life_table <- merge(input_life_table, select(male_ylds_interpolated ,c(sex_age_cat, pyld_rate)) , by = 'sex_age_cat', all.x = TRUE) 
input_life_table <- merge(input_life_table, select(female_ylds_interpolated ,c(sex_age_cat, pyld_rate)), by = 'sex_age_cat', all.x = TRUE)

@@

input_life_table$mx <-  ifelse(input_life_table$sex == "male",
input_life_table$mx.x, input_life_table$mx.y)
input_life_table$pyld_rate <-  ifelse(input_life_table$sex == "male",
input_life_table$pyld_rate.x, input_life_table$pyld_rate.y)
## Drop redundant variables
input_life_table <- subset(input_life_table, select = -c(mx.y, mx.x, pyld_rate.x, pyld_rate.y))

## Cross check population numbers
sum(input_life_table$five_year_population, na.rm = TRUE)
## Sort data by sex and age to use in life table function. 
input_life_table<-input_life_table[order(input_life_table$sex, input_life_table$age),]

@@

```{r, echo=TRUE}
## Use externaly generated inputs for for disease life table. The previous code will serve to to some of the data preparation for DISMOD. Also need to devlop code to calculated DWs, for now, all generated in excel.

@@

for (age in i_age_cohort){
  for (sex in i_sex){
    for (disease in i_disease) {
      # Exclude breast_cancer for Males
      if (sex == "male" && disease == "breast_cancer"){
       # cat("\n") #Uncomment to see list
      }
      else {
        # cat("age ", age, " sex ", sex, "and disease", disease, "\n") #Uncomment to see list
        disease_life_table_list_bl[[index]] <- run_disease(in_idata = idata, in_sex = sex, in_mid_age = age, in_disease = disease)
        index <- index + 1
      }
    }
  }
}
##### Uncommnet to check disease life table list
# View(disease_life_table_list_bl[[8]])

@@

#####Generate scenario incidence (for each disease)

@@

##### Uncommnet to check scenario incidence
# View(incidence_sc[[1]])
```

@@

        # Instead of idata, feed td to run scenarios
        disease_life_table_list_sc[[index]] <- run_disease(in_idata = td1, in_sex = sex, in_mid_age = age, in_disease = disease)
        disease_life_table_list_sc[[index]]$diff_inc_disease <- disease_life_table_list_sc[[index]]$incidence_disease - disease_life_table_list_bl[[index]]$incidence_disease
        disease_life_table_list_sc[[index]]$diff_prev_disease <- disease_life_table_list_sc[[index]]$px - disease_life_table_list_bl[[index]]$px
        disease_life_table_list_sc[[index]]$diff_mort_disease <- disease_life_table_list_sc[[index]]$mx - disease_life_table_list_bl[[index]]$mx
        disease_life_table_list_sc[[index]]$diff_pylds_disease <- (disease_life_table_list_sc[[index]]$px - disease_life_table_list_bl[[index]]$px) * disease_life_table_list_bl[[index]]$dw_disease
        index <- index + 1
      }
    }
  }
}
##### Uncommnet to check scenario life tables
# View(disease_life_table_list_sc[[1]])

@@

```{r, echo=TRUE}
#####Generate total change in mortality rate
##### Sum mortality rate scenarios (mx_sc_total)

@@

##### Uncommnet to check sceanrio mortality and changes 
# View(mx_sc_total[[1]])

#####Generate total change in prevalent yld rates
#####total ylds rate= sum (change prevalence disease*dw)

@@

##### Uncommnet to check scenario pyld change
# View(pylds_sc_total[[2]])

##### Calculate general life tables with modified mortality and pylds total#############################
###Original mortality rate is modified by the mx_sc_total (total change in mortality from diseases)
###Original pyld rate is modified by the change in each disease pylds

@@

    index <- index + 1
  }
}
##### Uncommnet to check scenario life tables
# View(general_life_table_list_sc[[1]])
# View(general_life_table_list_bl[[2]])
```

21) Generate list of outputs by age and sex

```{r, echo=TRUE}
#####In the following list "output_life_table", 32 data frames are nested per age and sex cohort

@@

    #Males do not have breast cancer, that is why we need the if/else. 
    #We create a TRUE/FALSE variable for the loop to move into the next disease

@@

          ####New list to add calculations for changes in burden of disease (incidence and mortality numbers)

@@

          ###Bind all lists

@@

          #Here the calculations above are repeated, here is where the F is telling to move into the next disease

@@

    # general_life_table_list_sc and general_life_table_list_bl (Lx)
    output_burden_lf_sc <- select(general_life_table_list_sc[[l_index]], c('Lx', 'Lwx'))
    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lx'] <- paste('Lx', "sc", sep = "_")
    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lwx'] <- paste('Lwx', "sc", sep = "_")

@@

#Uncomment to check
# View(output_burden[[32]])

@@

 p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, sep = " "))
p_list[[index]] <- p_index
index <- index + 1

@@

# m<-plot_grid(p_index, ncol=2, nrow=2)
# library(cowplot)
# save_plot("output.tiff",m ,ncol = 2, nrow = 2) # make room for figure legend)

# png(file="superplot.png")
# marrangeGrob(grobs=p_list, nrow=1, ncol=1)
# # do.call("grid.arrange", p_index)

ml <- do.call(marrangeGrob, list(grobs=p_list, nrow = 3, ncol = 2))
ml

>>>>

## Global options
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = 'latex')
knitr::opts_chunk$set(fig.cap = TRUE, fig.align = "center", fig.pos = "H", fig.scap = TRUE)

## The work in this document requires the following packages (code, rmarkdown and citing packages):

require(dplyr)
require(tidyverse)
require(knitr)
require(kableExtra)
require(citr)
require(gridExtra)
require(cowplot)
require(ggpubr)

@@

```{r, echo=FALSE}

pmslt_inputs <- tribble(
~Input, ~Source, ~Comments, 
"Life table", "Synthetic population per sex and age group", "Age grouping in life table to match synthetic population", 
"Life table", "Synthetic population per sex and one-year age group", "If one year age group is not avabilable it can be derive using interpolation from age groups data", 
"Life table", "Global Burden of Disease (GBD) study per one-year age group and sex", "GBD data is in five-year age groups, interpolation to derive one-year age groups", 
"Disease life table", "GBD data for prevalence, incidence and mortality and DISMOD II", "Two step process. First obtain disease and population data from GBD. Second, use Dismod II to derive internally consistent estimates for incidence and case fatality (PMSLT disease life table iputs)", 
"Disease life table", "Derive from disease prevalence and years lived with disability from GBD", "Adjustments for comorbidities in later years of life to be applied"
)

@@

```{r, echo=FALSE}

pmslt_inputs <- tribble(
~Tab, ~Selection,
"Base", "Single", "Location", "Case study city", "Year", "Latest available", "Context", "Cause", "Age", "Under 5, 5 to 9, 10 to 14, 15 to 19, 20 to 24, 25 to 29, 30 to 34, 35 to 39, 40 to 49, 50 to 54, 55 to 59, 60 to 64, 65 to 69, 70 to 74, 75 to 79, 80 to 84, 89 to 89, 90 to 95, 95 plus", "Metric", "Number, Rate", "Measure", "Deaths, YLDs, Prevalence, Incidence", "Sex", "Male, Female", "Cause", "Total All causes, ischemic heart disease, etc")

kable(pmslt_inputs,"latex", booktabs = T, caption = "Global burden of disease data") %>%
 kable_styling(latex_options = c("striped", "hold position"),
               full_width = F)%>%
  column_spec(2, width = "30em") %>%
  column_spec(3, width = "30em") %>%
  row_spec(0, bold = T)

@@

4) Read GBD data
```{r, echo=TRUE}

GBDdata <- read.csv("data/UK/gbd2016.csv", stringsAsFactors = F)

@@

5) Change all upper cases to lower case and delete () from variables. 
```{r, echo=TRUE}

GBDdata <- mutate_all(GBDdata, funs(tolower))
GBDdata$measure[GBDdata$measure=="ylds (years lived with disability)"] <- "ylds"

6) Create age categories index in GBDdata (mid age, to match cohort running), total of 20 age groups. These are the age cohorts to simulate. 
```{r, echo=TRUE}

# 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97
GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 7
GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 12
GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 17
GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 22
GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 27
GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 32
GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 37
GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 42
GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 47
GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 52
GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 57
GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 62
GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 67
GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 72
GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 77
GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 82
GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 87
GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 92
GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97

7) Create age and sex categories to obtain population numbers. Population numbers from GBD are used in Dismod II. For the Life table (Figure 3), the numbers may be from the synthetic population. For now, the Life table is set up with population numbers derived from the GBD data. 
```{r, echo=TRUE}

GBDdata$sex_age_cat <- paste(GBDdata$sex,GBDdata$age_cat, sep = "_"  )

```
8) Convert string variables to numeric to do calculations.
```{r, echo=TRUE}

GBDdata$val <- as.numeric(as.character(GBDdata$val))

```
9) Generate population numbers for Greater London in a new data frame ("GBD_population"). Note that there is data for England as well, which is used in a separate rmarkdown document (GBDCompare). Population numbers are derived from rates per 100,000 and total numbers of cases. 
```{r, echo=TRUE}

GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", 
metric == "rate" | metric == "number" ) %>% select(metric, age_cat, sex, val, 
sex_age_cat, location)

```
10) Generate population numbers from given number of cases and rates per 100,000 people. 
```{r, echo=TRUE}

 for (i in 1:nrow(GBD_population)) {
    if (GBD_population$metric[i] == "number") {
      GBD_population$five_year_population[i] <- GBD_population$val[i] * 100000/
      GBD_population$val[i + 2]}     
    else {GBD_population$five_year_population[i] <- NA}
 }

```
11) Remove rows with zero
```{r, echo=TRUE}

GBD_population <- GBD_population[!is.na(GBD_population$five_year_population),]

```
12) Keep relevant variables
```{r, echo=TRUE}

GBD_population <- filter(GBD_population) %>% 
select(sex_age_cat, sex, age_cat, five_year_population, location)

```
13) Create data frames for Greater London to be later used for: a) interpolation of rates and b) PMLT cohorts. 
```{r, echo=TRUE}

GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% 
select(sex_age_cat, age_cat, sex, five_year_population, location, sex_age_cat)

```
12) Check population total numbers
```{r, echo=TRUE}

GreaterLondon <- sum(GBD_population_GL$five_year_population)

13) Generate data frames for Greater London with per person rates (per 100,000 in original data).  
```{r, echo=TRUE}

GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% 
select(measure, location, sex, age, metric, cause, val, age_cat, sex_age_cat)
GBDGL$one_rate <- GBDGL$val/100000

@@

# Name data frame
assign(paste(sex_index, measure_index, "interpolated", sep = "_"), interpolated)

## Do graph with another layer for original rates in age groups for comparison purposes.

p_interpolation_list <- list()

interpolation_index <- 1

p_interpolation_index <- ggplot(data = interpolated,mapping = aes(age, interpolated[,1])) +
              geom_line(aes(color = "Interpolated")) +
              geom_point(
              data = data, 
              mapping = aes(age_cat, one_rate, color = "Original")) +
              labs(colour="",x="Age",y="Rates", sep = " ") +
              labs (title = paste("Rates", sex_index, "all cause", 
              measure_index, sep = " "), size=14) +
              theme_classic() +
              theme (plot.title = element_text(hjust = 0.5))

              p_interpolation_list[[interpolation_index]] <- p_interpolation_index
              interpolation_index <- interpolation_index + 1

## Save plots to tiff. Makes a separate file for each plot.

interpolation_index <- 1
for(sex_index in i_sex) {
  for (measure_index in i_measure) {
    file_name = paste("graphs/", "Interpolation rates", sex_index, measure_index, ".tiff", sep=" ")
    tiff(file_name)
    print(p_interpolation_list[[interpolation_index]])
    interpolation_index <- interpolation_index + 1
    dev.off()
  }
}

p_interpolated <- do.call(marrangeGrob, list(grobs=p_interpolation_list, nrow = 2, ncol = 2))
p_interpolated

@@

## Create empty data frame

input_life_table <- as.data.frame(matrix(0, ncol=2, nrow = 202))

## Give variables names (mx=total mortality rate, 
## yldsx = total years lived with disability rates )

names(input_life_table) <- c("age", "sex")

## Populate life table: Age, 0 to 100

input_life_table[1:101, 1] <- c(0:100)
input_life_table[102:202, 1] <- c(0:100)

## Populate life table: sex, male, female

input_life_table[1:101, 2] <- "male"
input_life_table[102:202, 2] <- "female"

## Create variable age_sex to match with population data

input_life_table$sex_age_cat <- paste(input_life_table$sex,input_life_table$age,
                                sep = "_"  )
## Populate life table: five_year_pop.

input_life_table <- merge(input_life_table, select(GBD_population_GL, 
                    c(sex_age_cat, five_year_population)), by = 'sex_age_cat', all = TRUE)

## Populate life table: mortality rates (from interpolated rates)
## Males mortality rates. NEED MATCHING NAMES

input_life_table <- merge(input_life_table, select(male_deaths_interpolated, 
                    c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
input_life_table <- merge(input_life_table, select(female_deaths_interpolated,
                    c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
input_life_table <- merge(input_life_table, select(male_ylds_interpolated, 
                    c(sex_age_cat, pyld_rate)) , by = 'sex_age_cat', all.x = TRUE) 
input_life_table <- merge(input_life_table, select(female_ylds_interpolated, 
                    c(sex_age_cat, pyld_rate)), by = 'sex_age_cat', all.x = TRUE)

@@

input_life_table$mx <-  ifelse(input_life_table$sex == "male",
                        input_life_table$mx.x, input_life_table$mx.y)

input_life_table$pyld_rate <-  ifelse(input_life_table$sex == "male",
                              input_life_table$pyld_rate.x, input_life_table$pyld_rate.y)
## Drop redundant variables
input_life_table <- subset(input_life_table, select = -c(mx.y, mx.x, pyld_rate.x, pyld_rate.y))

## Cross check population numbers

sum(input_life_table$five_year_population, na.rm = TRUE)

## Sort data by sex and age to use in life table function

input_life_table<-input_life_table[order(input_life_table$sex, input_life_table$age),]

@@

## Use externaly generated inputs for for disease life table. The previous code will serve to to some of the data preparation for DISMOD. Also need to devlop code to calculated DWs, for now, all generated in excel.

@@

for (age in i_age_cohort){
  for (sex in i_sex){
    for (disease in i_disease) {
      # Exclude breast_cancer for Males
      if (sex == "male" && disease == "breast_cancer"){
       # cat("\n") #Uncomment to see list
      }
      else {
        # cat("age ", age, " sex ", sex, "and disease", disease, "\n") #Uncomment to see list
        disease_life_table_list_bl[[index]] <- run_disease(in_idata = idata, in_sex = sex, in_mid_age = age, 
                                               in_disease = disease)
        index <- index + 1
      }
    }
  }
}
## Uncommnet to check disease life table list
# View(disease_life_table_list_bl[[8]])

@@

## Generate scenario incidence (for each disease)

@@

## Uncommnet to check scenario incidence
# View(incidence_sc[[1]])
```

@@

        # Instead of idata, feed td to run scenarios
        disease_life_table_list_sc[[index]] <- run_disease(in_idata = td1, in_sex = sex, 
                                               in_mid_age = age, in_disease = disease)
        disease_life_table_list_sc[[index]]$diff_inc_disease <- 
                                              disease_life_table_list_sc[[index]]$incidence_disease 
                                               -   disease_life_table_list_bl[[index]]$incidence_disease
        disease_life_table_list_sc[[index]]$diff_prev_disease <- 
                                              disease_life_table_list_sc[[index]]$px 
                                              - disease_life_table_list_bl[[index]]$px
        disease_life_table_list_sc[[index]]$diff_mort_disease <- 
                                              disease_life_table_list_sc[[index]]$mx 
                                              - disease_life_table_list_bl[[index]]$mx
        disease_life_table_list_sc[[index]]$diff_pylds_disease <- 
                                              (disease_life_table_list_sc[[index]]$px  
                                              - disease_life_table_list_bl[[index]]$px) * 
                                              disease_life_table_list_bl[[index]]$dw_disease
        index <- index + 1
      }
    }
  }
}
## Uncommnet to check scenario life tables
# View(disease_life_table_list_sc[[1]])

@@

## Generate total change in mortality rate
## Sum mortality rate scenarios (mx_sc_total)

@@

## Uncommnet to check sceanrio mortality and changes 
# View(mx_sc_total[[1]])

## Generate total change in prevalent yld rates
## Total ylds rate= sum (change prevalence disease*dw)

@@

## Uncommnet to check scenario pyld change
# View(pylds_sc_total[[2]])

## Calculate general life tables with modified mortality and pylds total
## Original mortality rate is modified by the mx_sc_total (total change in mortality from diseases)
## Original pyld rate is modified by the change in each disease pylds

@@

    index <- index + 1
  }
}
## Uncommnet to check scenario life tables
# View(general_life_table_list_sc[[1]])
# View(general_life_table_list_bl[[2]])

21) Generate list of outputs by age and sex.

## In the following list "output_life_table", 32 data frames are nested per age and sex cohort

@@

    # Males do not have breast cancer, that is why we need the if/else. 
    # We create a TRUE/FALSE variable for the loop to move into the next disease

@@

          ## New list to add calculations for changes in burden of disease (incidence and mortality numbers)

@@

          ## Bind all lists

@@

          ## Here the calculations above are repeated, here is where the F is telling to move into the next disease

@@

    ## general_life_table_list_sc and general_life_table_list_bl (Lx)
    output_burden_lf_sc <- select(general_life_table_list_sc[[l_index]], 
                          c('Lx', 'Lwx'))
    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lx'] <- 
                          paste('Lx', "sc", sep = "_")
    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lwx'] <- 
                          paste('Lwx', "sc", sep = "_")

@@

## Uncomment to check
# View(output_burden[[32]])

@@

p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, sep = " "))

p_list[[index]] <- p_index
index <- index + 1

@@

## Use plot_grid in a loo/grid.arrange

mpage <- do.call(marrangeGrob, list(grobs=p_list, nrow = 3, ncol = 2))
mpage

## Trying to use ggarrange
# mpage2 <- ggarrange(plotlist = p_list, ncol = 3, nrow = 3)

# ggsave("graphs/multipage.jpeg", mpage)
